`Big data' og politologisk datavidenskab

### 2020-03-21 draft_v01.Rmd
update 'hvad er big data' section w example

The hurdle of accessing big data underscores a basic difference between it and traditional social science data: traditional social science data were collected for the purpose of doing social science. Social scientists must always consider the sources and the nature of our data, and big data has sometimes been referred to as "found data" or "digital exhaust" [@harford14]. What does this mean? Virtually all big data are purpose-built for goals *other than* social research. Metadata like timestamps, follower counts, or activity frequencies on social media web sites are not stored for science or, necessarily, according to scientific standards. Although they may be useful, their scientific value is an unintended byproduct (i.e. exhaust) of business or government activity. As such, when applying big data to social science research we must always probe the implications of the data's purpose for our scientific applications.^[As @salganik17 puts it, the challenges and opportunities created by big data follow from asking why the data were collected.]

- How/why big data became what it is in the zeitgeist
  - Classic advantages: Big / always on
  - Classic disadvantages: metered/restricted access + expertise barrier

- Definition of machine learning: "y-hat vs. beta-hat"

@@

Systems that collect big data are purpose-built, and the purpose is never political science research. This holds true even for the most research-friendly data sources.

<!-- Include example about retsinformation.dk?? Ministry titles in official metadata reflect "current" institutional structure and throws out information contemporary to older data. This creates a massive headache for researchers, but it makes perfect sense for running an efficient bookkeeping operation for the central bureaucracy. Furthermore, there are not metadata linking laws to the decrees, regulations, etc. that they enable. This system would be superfluous and redundant since retsinformation is a broad-based clearing house and wouldn't necessarily be the tool of choice for legal advisors drafing secondary legislation within the line ministries. The data in the official gazette are complete, but they come with major biases caused directly by the fact that--even this public, non-partisan resource--is built with a purpose that is different from (not necessarily opposed to) political science research.

To cap it off, retsinformation.dk isn't finished evolving. It will continue to develop in ways that may or may not benefit political science research as its maintainers' and users' needs and preferences change.-->

- Big data typically *not designed for research*, i.e.
  - it's dirty (re: Salganik) -- human behavior is mixed together with actions taken by bots/automated systems
  - drifting -- Needs of *actual* system maintainers / users may be completely orthogonal to needs of political science research (or even opposed, consider FB)
  - algorithmically confounded -- large-scale systems have robot nannies. These include everything from spell checkers to YouTube's recommendation algorithm. The observed behavior we find in big data is a consequence of the (generally) unobservable interaction between humans and these algorithms.

- Standard sampling issues (nonrepresentative, systematic sampling bias)

>>>>

This highlights an important difference between big data and traditional social science data: traditional social science data were collected for the purpose of social science. Big data, by contrast, is sometimes referred to as "found data" or "digital exhaust" [@harford14]. What does this mean? Virtually all big data are purpose-built for goals *other than* social research. Metadata like timestamps, follower counts, or activity frequencies on social media web sites are not stored for science or, necessarily, according to scientific standards.

Even government data sources present this problem, sometimes in confusing ways. Consider a minor example: lov nr 1049 af 11/12/1996. The law is short, fewer than 20 words. The only thing it does is ophæve lov om skoleskibsafgift. Civilstyrelsens database of all Danish law, *retsinformation.dk*, lists several pieces of metadata about lov nr 1049.^[See: https://www.retsinformation.dk/Forms/R0710.aspx?id=83509] For example, the metadata capture the law it repeals, several related documents, its publication date, and ministerområdet. This last one is confusing, however. Ministerområdet for lov nr 1049 is currently listed as Uddannelses- og Forskningsministeriet. However, the law was signed by Mimi Jakobsen. If you do some further research, you will find that Mimi Jakobsen var Erhvervsminister i Regeringen Poul Nyrup Rasmussen II.^[https://www.regeringen.dk/regeringer-siden-1848/regeringen-poul-nyrup-rasmussen-ii/] Why, then, is this law not connected to Erhvervsministeriet? This confusion is caused because Civilstyrelsen opdaterer ministerområder løbende based on *today's* legal framework and ministries. Connecting lov nr 1049 to Uddannelses- og Forskningsministeriet is obviously wrong if you want to build a historical data set of Danish law. Unfortunately for political scientists, it is correct if---like Civilstyrelsen---your goal is not historical research, but instead to organize the legislation currently in force in Denmark. To make matters worse for researchers, retsinformation.dk will continue to evolve in ways that may or may not benefit political science research, depending on Civilstyrelsens future needs and goals.

When using big data built for some official or private purpose, problems of this nature are always lurking. Although big data can be extremely useful for social science, their scientific value is an *unintended byproduct* (i.e. exhaust) of business or government activity. When applying big data to social science research we must always understand why and how the data were assembled and probe the implications of the data's purpose for our scientific applications.^[As @salganik17 puts it, the challenges and opportunities created by big data follow from asking why the data were collected.]

 <!-- - Definition of machine learning: "y-hat vs. beta-hat" -->

@@

<!-- Systems that collect big data are purpose-built, and the purpose is never political science research. This holds true even for the most research-friendly data sources. -->

Big data in political science research designs:
  - addressing confounding
    - it's dirty (re: Salganik) -- human behavior is mixed together with actions taken by bots/automated systems
    - drifting -- Needs of *actual* system maintainers / users may be completely orthogonal to needs of political science research (or even opposed, consider FB)
    - algorithmically confounded -- large-scale systems have robot nannies. These include everything from spell checkers to YouTube's recommendation algorithm. The observed behavior we find in big data is a consequence of the (generally) unobservable interaction between humans and these algorithms.
    - Standard sampling issues (nonrepresentative, systematic sampling bias)
  - its role
    - uncommon to directly analyze
    - more often: measurement

### 2020-03-21 draft_v01.tex
update 'hvad er big data' section w example

\documentclass[12pt,]{article}
\usepackage[]{cochineal}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={`Big data' og politologisk datavidenskab},
            pdfauthor={Frederik Hjorth; Matt W. Loftis},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{apsr}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{`Big data' og politologisk datavidenskab}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \subtitle{Udkast, april 2020}
  \author{Frederik Hjorth \\ Matt W. Loftis}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

@@

\hypertarget{hvad-er-big-data}{%
\subsection{Hvad er big data?}\label{hvad-er-big-data}}

Big data has been credited with enabling an emerging field of
computational social science with ``the capacity to collect and analyze
data with an unprecedented breadth and depth and scale''
\citep{lazeretal}. Yet, the concept of big data lives two lives. On the
one hand the popular definition is associated with exciting and
futuristic promises of data-driven science and technology. On the other
hand is what we might call the operational definition of how massive
data sets are collected, stored, and used currently by social
scientists. To unpack the role of big data in political science
research, we first consider the second aspect of the concept and then
circle back to examine what to believe about the first aspect.

One widely familiar operational definition identifies big data with the
so-called three Vs: Volume, Variety, and Velocity \citep{laney01}. Big
data, here, refers to fast-evolving and very large data sets--including
even data too large to store on a desktop computer--that contain lots of
variation. Data sources of this nature are often interesting to social
science: search engine logs, social media activity, government
administrative records, mobile telephone records, or even data stored by
passive monitoring conducted by digital devices in the physical world
have been mobilized for social science research \citep{salganik17}.
Accessing these data requires partnerships with their owners--phone
companies, governments, technology companies, etc. This can mean writing
software to access websites or remote databases or can involve more
formalized partnerships to share information securely \citep{EL14}.

The hurdle of accessing big data underscores a basic difference between
it and traditional social science data: traditional social science data
were collected for the purpose of doing social science. Social
scientists must always consider the sources and the nature of our data,
and big data has sometimes been referred to as ``found data'' or
``digital exhaust'' \citep{harford14}. What does this mean? Virtually
all big data are purpose-built for goals \emph{other than} social
research. Metadata like timestamps, follower counts, or activity
frequencies on social media web sites are not stored for science or,
necessarily, according to scientific standards. Although they may be
useful, their scientific value is an unintended byproduct (i.e.~exhaust)
of business or government activity. As such, when applying big data to
social science research we must always probe the implications of the
data's purpose for our scientific applications.\footnote{As
  \citet{salganik17} puts it, the challenges and opportunities created
  by big data follow from asking why the data were collected.}

\begin{itemize}
\tightlist
\item
  How/why big data became what it is in the zeitgeist

  \begin{itemize}
  \tightlist
  \item
    Classic advantages: Big / always on
  \item
    Classic disadvantages: metered/restricted access + expertise barrier
  \end{itemize}
\end{itemize}

\hypertarget{big-data-og-forskningsdesign}{%
\subsection{Big data og
forskningsdesign}\label{big-data-og-forskningsdesign}}

Systems that collect big data are purpose-built, and the purpose is
never political science research. This holds true even for the most
research-friendly data sources.

\begin{itemize}
\tightlist
\item
  Big data typically \emph{not designed for research}, i.e.

  \begin{itemize}
  \tightlist
  \item
    it's dirty (re: Salganik) -- human behavior is mixed together with
    actions taken by bots/automated systems
  \item
    drifting -- Needs of \emph{actual} system maintainers / users may be
    completely orthogonal to needs of political science research (or
    even opposed, consider FB)
  \item
    algorithmically confounded -- large-scale systems have robot
    nannies. These include everything from spell checkers to YouTube's
    recommendation algorithm. The observed behavior we find in big data
    is a consequence of the (generally) unobservable interaction between
    humans and these algorithms.
  \end{itemize}
\item
  Standard sampling issues (nonrepresentative, systematic sampling bias)
\end{itemize}

\hypertarget{behandling-af-big-data}{%
\subsection{Behandling af big data}\label{behandling-af-big-data}}

@@

  \begin{itemize}
  \tightlist
  \item
    Generic data: Clustering, IRT, etc.
  \item
    Text: Topic modeling, text scaling, dictionaries
  \item
    Fundamental unity of goals and approach
  \item
    Variety in methods results from variation in:

    \begin{itemize}
    \tightlist
    \item
      Assumptions re: underlying model/geometry of latent space
    \item
      Related to above: something like, ``format'' of output
    \item
      Structure/nature of input data
    \item
      Amount of domain expertise applied to structure results
    \item
      Assumptions about what is correlated with what
    \item
      Level of computational intensity
    \end{itemize}
  \end{itemize}
\item
  Classification / prediction
\item
  Explanatory modeling
\end{itemize}

\hypertarget{etiske-problemer-ved-big-data}{%
\subsection{Etiske problemer ved big
data}\label{etiske-problemer-ved-big-data}}

>>>>

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[12pt,]{article}
\usepackage[]{cochineal}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={`Big data' og politologisk datavidenskab},
            pdfauthor={Frederik Hjorth; Matt W. Loftis},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\usepackage[]{natbib}
\bibliographystyle{apsr}

\title{`Big data' og politologisk datavidenskab}
\providecommand{\subtitle}[1]{}
\subtitle{Udkast, april 2020}
\author{Frederik Hjorth \and Matt W. Loftis}
\date{}

@@

\hypertarget{hvad-er-big-data}{%
\section{Hvad er big data?}\label{hvad-er-big-data}}

I en toneangivende artikel peger \citet{lazeretal} på big data som
kilden til en ny type samfundsvidenskab, ``computational social
science'', med ``kapacitet til at indsamle og analysere data med
historisk uset bredde, dybde og omfang''. Men begrebet big data lever to
liv. På den ene side er den populære definition af begrebet forbundet
med futuristiske løfter om ny, datadrevet videnskab og teknologi. På den
anden side står hvad man kunne kalde den operationelle definition af
hvordan store datamænger indsamles, lagres og analyseres af
samfundsvidenskabsfolk. For at belyse betydningen af big data for
politologi betragter vi først denne anden, operationelle betydning af
begrebet inden vi vender tilbage til den første, populære betydning.

En bredt anvendt operationel definition identificerer big data med de
såkaldte `tre V'er': \emph{Volume}, \emph{Variety} og \emph{Velocity},
dvs. volumen, variation og hastighed \citep{laney01}. Big data refererer
her til foranderlige og meget store datasæt -- inklusive data der er for
store til at gemme på en almindelig pc -- der indeholder masser af
variation. Datakilder af denne art er ofte interessante for
samfundsvidenskab: søgemaskinelogfiler, sociale medieaktiviteter,
offentlige registre, mobiltelefonregistre eller endda data gemt ved
passiv overvågning udført af digitale enheder i den fysiske verden er
alle blevet anvendt til samfundsvidenskabelig forskning
\citep{salganik17}. Adgang til disse data kræver partnerskaber med deres
ejere -- telefonfirmaer, regeringer, teknologiselskaber osv. Dette kan
betyde, at man skriver software til at få adgang til websteder eller
eksterne databaser eller kan involvere mere formaliserede partnerskaber
for at dele information sikkert \citep{EL14}. I afsnittet om
``Anskaffelse af data'' nedenfor beskriver vi nærmere hvordan det kan
finde sted.

This highlights an important difference between big data and traditional
social science data: traditional social science data were collected for
the purpose of social science. Big data, by contrast, is sometimes
referred to as ``found data'' or ``digital exhaust'' \citep{harford14}.
What does this mean? Virtually all big data are purpose-built for goals
\emph{other than} social research. Metadata like timestamps, follower
counts, or activity frequencies on social media web sites are not stored
for science or, necessarily, according to scientific standards.

Even government data sources present this problem, sometimes in
confusing ways. Consider a minor example: lov nr 1049 af 11/12/1996. The
law is short, fewer than 20 words. The only thing it does is ophæve lov
om skoleskibsafgift. Civilstyrelsens database of all Danish law,
\emph{retsinformation.dk}, lists several pieces of metadata about lov nr
1049.\footnote{See:
  \url{https://www.retsinformation.dk/Forms/R0710.aspx?id=83509}} For
example, the metadata capture the law it repeals, several related
documents, its publication date, and ministerområdet. This last one is
confusing, however. Ministerområdet for lov nr 1049 is currently listed
as Uddannelses- og Forskningsministeriet. However, the law was signed by
Mimi Jakobsen. If you do some further research, you will find that Mimi
Jakobsen var Erhvervsminister i Regeringen Poul Nyrup Rasmussen
II.\footnote{\url{https://www.regeringen.dk/regeringer-siden-1848/regeringen-poul-nyrup-rasmussen-ii/}}
Why, then, is this law not connected to Erhvervsministeriet? This
confusion is caused because Civilstyrelsen opdaterer ministerområder
løbende based on \emph{today's} legal framework and ministries.
Connecting lov nr 1049 to Uddannelses- og Forskningsministeriet is
obviously wrong if you want to build a historical data set of Danish
law. Unfortunately for political scientists, it is correct if---like
Civilstyrelsen---your goal is not historical research, but instead to
organize the legislation currently in force in Denmark. To make matters
worse for researchers, retsinformation.dk will continue to evolve in
ways that may or may not benefit political science research, depending
on Civilstyrelsens future needs and goals.

When using big data built for some official or private purpose, problems
of this nature are always lurking. Although big data can be extremely
useful for social science, their scientific value is an \emph{unintended
byproduct} (i.e.~exhaust) of business or government activity. When
applying big data to social science research we must always understand
why and how the data were assembled and probe the implications of the
data's purpose for our scientific applications.\footnote{As
  \citet{salganik17} puts it, the challenges and opportunities created
  by big data follow from asking why the data were collected.}

\hypertarget{big-data-og-forskningsdesign}{%
\section{Big data og
forskningsdesign}\label{big-data-og-forskningsdesign}}

Big data in political science research designs: - addressing confounding
- it's dirty (re: Salganik) -- human behavior is mixed together with
actions taken by bots/automated systems - drifting -- Needs of
\emph{actual} system maintainers / users may be completely orthogonal to
needs of political science research (or even opposed, consider FB) -
algorithmically confounded -- large-scale systems have robot nannies.
These include everything from spell checkers to YouTube's recommendation
algorithm. The observed behavior we find in big data is a consequence of
the (generally) unobservable interaction between humans and these
algorithms. - Standard sampling issues (nonrepresentative, systematic
sampling bias) - its role - uncommon to directly analyze - more often:
measurement

\hypertarget{behandling-af-big-data}{%
\section{Behandling af big data}\label{behandling-af-big-data}}

\hypertarget{anskaffelse-af-data}{%
\subsection{Anskaffelse af data}\label{anskaffelse-af-data}}

\begin{itemize}
\tightlist
\item
  Getting data: scraping, APIs, text databases
\end{itemize}

\hypertarget{usuperviserede-tilgange}{%
\subsection{Usuperviserede tilgange}\label{usuperviserede-tilgange}}

\hypertarget{superviserede-tilgange}{%
\subsection{Superviserede tilgange}\label{superviserede-tilgange}}

\hypertarget{tekst-som-data}{%
\subsection{Tekst som data}\label{tekst-som-data}}

@@

  \begin{itemize}
  \tightlist
  \item
    Generic data: Clustering, IRT, etc.
  \item
    Fundamental unity of goals and approach
  \item
    Variety in methods results from variation in:

    \begin{itemize}
    \tightlist
    \item
      Assumptions re: underlying model/geometry of latent space
    \item
      Related to above: something like, ``format'' of output
    \item
      Structure/nature of input data
    \item
      Amount of domain expertise applied to structure results
    \item
      Assumptions about what is correlated with what
    \item
      Level of computational intensity
    \end{itemize}
  \end{itemize}
\item
  Text: Topic modeling, text scaling, dictionaries
\item
  Classification / prediction
\item
  Explanatory modeling
\end{itemize}

\hypertarget{etiske-problemer-ved-big-data}{%
\section{Etiske problemer ved big
data}\label{etiske-problemer-ved-big-data}}

emotional contagion example: \citet{kramer2014experimental}

problem: lack of informed consent

complication: under surveillance capitalism, all citizens are subject to
constant experimentation w/o consent
